Long term cloud enhancement events of global solar irradiance over Thessaloniki

### 2024-06-03 article/article.tex
Commit 2024-06-03 13:10

\documentclass[preprint, 3p,
authoryear]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%

@@

  \journal{An awesome journal} % Sets Journal name

@@

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
% TODO: Currently lineno needs to be loaded after amsmath because of conflict
% https://github.com/latex-lineno/lineno/issues/5
\usepackage{lineno} % add
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}

@@

  \title{Long term cloud enhancement events of global solar irradiance
over Thessaloniki}
    \author[LAP]{Athanasios N. Natsis%
  \corref{cor1}%
  \fnref{1}}
   \ead{natsisphysicist@gmail.com} 
    \author[LAP]{Alkiviadis Bais%
  %
  }
   \ead{abais@auth.gr} 
    \author[LAP]{Charikleia Meleti%
  %
  \fnref{2}}
   \ead{meleti@auth.gr} 
      \affiliation[LAP]{
    organization={Laboratory of Atmospheric Physics, Physics Department,
Aristotle University of
Thessaloniki},city={Thessaloniki},postcode={54124},country={Greece},}
    \affiliation[Another University]{
    organization={Department},addressline={A street
29},city={Manchester,},postcode={2054 NX},state={State},country={The
Netherlands},}
    \cortext[cor1]{Corresponding author}
    \fntext[1]{This is the first author footnote.}
    \fntext[2]{Another author footnote.}

  \begin{abstract}
  Here wii be the abstract.

  It consists of two paragraphs.
  \end{abstract}
    \begin{keyword}
    cloud enhancement \sep total solar radiation \sep global horizonta
irradiance \sep 
    over irradiance
  \end{keyword}

@@

~ ~ CE: cloud enhancement cases one minute

@@

{\centering \includegraphics[width=0.5\linewidth]{../images/P_CS_change-1}

@@

Our dataset, after the data selection processing, consists of 6144534
records of GHI, of which \(1.764\,\%\) are CE events and \(0.036\,\%\)
are ECE events. The highest GHI recorder was \(1416.6\,W/m^2\) on 24~May
2007. The absolute stronger CE event had an OIR of \(345.86\,W/m^2\) on
24~May 2007. The relative stronger CE event was \(49.7\,\%\) above the
clear sky threshold, on 28~October 2016.

\hypertarget{trends}{%
\subsection{Trends}\label{trends}}

@@

{\centering \includegraphics[width=0.5\linewidth]{../images/P_energy-6}

@@

\caption{Trend of yearly CE number of occurancies.}\label{fig:P-energy-N}
\end{figure}

@@

{\centering \includegraphics[width=0.5\linewidth]{../images/P_energy-5}

@@

\hypertarget{ce-climatology}{%
\subsection{CE Climatology}\label{ce-climatology}}

Another interesting aspect of the CE cases occurrences, is their
seasonal cycle. In
Figure\nobreakspace{}\ref{fig:relative-month-occurancies}, we have a
monthly box plot (whisker plot), where the values have been normalized
by the highest median value, that occurs on June. Although the number of
occurrences has a wide spread throughout the study period, the most
active period of CE occurrences is during May and June. During the
Winter (December -- February) the CE cases are about \(25\,\%\) of the
maximum. For the rest months the occurrences gradually ramp between the
maximum and minimum. These seasonality is a combinations of the clouds
occurrences, and the relevant position of the sun during the year.
Unfortunately, the lack of detailed data of cloud formation, type and
location is not allowing further analysis.

@@

{\centering \includegraphics[width=0.5\linewidth]{../images/clim_CE_month_norm_MAX_median_N-2}

@@

\caption{Seasonal statistics of the number of CE occurancies for each month, normalized to the maximum occurancies on June. The box represents the values of the low $25\,\%$ percentile to $75\,\%$ percentile, where the thick horizontal line inside is the mean, the verical lines extend to the macimum and minimum vales, the dots are outlier values, and the rhombus is the mean.}\label{fig:relative-month-occurancies}
\end{figure}

The distribution of the CE over irradiance intensity follows an
exponential decline Figure\nobreakspace{}\ref{fig:ovir-distribution},
where there is an inverse relation between the events frequency and
events intensity. This is expected as, the stronger the CE events
become, the rarer the particular atmospheric and sun conditions will
occur. This fact is indicative to the magnitude and probability of the
expected OIRs events over Thessaloniki. Similar distribution of CE
occurrences have been reported by \citet{Vamvakas2020}, where the
magnitude of OIR is higher due to the location of the the city of
Patras, \(2.5^\circ\) closer to the equator.

@@

{\centering \includegraphics[width=0.5\linewidth]{../images/P-relative-distribution-diff-1}

@@

{\centering \includegraphics[width=0.5\linewidth]{../images/groups-1}

@@

{\centering \includegraphics[width=0.5\linewidth]{../images/P-groups-bin2d-1}

@@

\caption{Relation of mean over irradiance and CE group duration}\label{fig:group-2d}
\end{figure}

\hypertarget{extreme-ce-above-tsi}{%
\subsection{Extreme CE above TSI}\label{extreme-ce-above-tsi}}

A characteristic of the CE events that is commonly reported and has some
significance on the solar energy production infrastructure are the
extreme CE events (ECE). Where solar irradiance exceeds the expected
irradiance on top of the atmosphere, we have defined this as ECE
(Equation\nobreakspace{}\ref{eq:ECE}).

Analogous to Figure\nobreakspace{}\ref{fig:relative-month-occurancies}
we have computed the distribution of the number of occurrences of ECE
events by month in
Figure\nobreakspace{}\ref{fig:relative-month-occurancies-ECE}. The most
active period for ECE events is in the spring and the start of the
summer (March -- June), followed by a period in the late fall (September
and October). This is probably an effect of the weather characteristics
of those periods where, there are continuous alternations between clear
sky periods and clouds.

@@

{\centering \includegraphics[width=0.5\linewidth]{../images/clim_ECE_month_norm_MAX_median_N-2}

@@

\caption{Seasonal statistics of the number of ECE occurancies for each month, normalized to the maximum occurancies on March. The box represents the values of the low $25\,\%$ percentile to $75\,\%$ percentile, where the thick horizontal line inside is the mean, the verical lines extend to the macimum and minimum vales, the dots are outlier values, and the rhombus is the mean.}\label{fig:relative-month-occurancies-ECE}
\end{figure}

The distribution of the ECE events
(Figure\nobreakspace{}\ref{fig:P-extreme-distribution}), shows that
there are rare cases where the OIR can exceed the TSI even more than
\(400\,\text{W}/\text{m}^2\), with the \(75\,\%\) of the cases to be
below \(200\,\text{W}/\text{m}^2\). Similar to CE cases, the
distribution of ECE have a similar pattern with the results of
\citet{Vamvakas2020}, with the difference that the OIR values are lower
fro Thessaloniki.

@@

{\centering \includegraphics[width=0.5\linewidth]{../images/P-extreme-distribution-1}

@@

{\centering \includegraphics[width=0.5\linewidth]{../images/example_days-18}

@@

\caption{Example of cloud indentification data for 2019-07-11. Where are denoted with red circles all the detected CE events, the cloud intearaction with the GHI (blue crosses), and the thesholds of CE and the threshold of ECE events with red and black line respectively.}\label{fig:example-day}
\end{figure}

{\centering \includegraphics[width=0.5\linewidth]{../images/example_years-13}

@@

\caption{Example of the relation of the measured GHI versus the reference clear sky irradiance for the year 2005, where the intesity of the detected over irradianse (OIR) is color coded.}\label{fig:example-year}
\end{figure}

>>>>

\documentclass[preprint, 5p,
authoryear]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%

@@

  \journal{Atmospheric Research} % Sets Journal name

@@

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
% TODO: Currently lineno needs to be loaded after amsmath because of conflict
% https://github.com/latex-lineno/lineno/issues/5
\usepackage{lineno} % add
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num-names}

@@

  \title{Long term cloud enhancement events of global solar irradiance
over Thessaloniki}
    \author[LAP]{Athanasios N. Natsis%
  \corref{cor1}%
  }
   \ead{natsisphysicist@gmail.com} 
    \author[LAP]{Alkiviadis Bais%
  %
  }
   \ead{abais@auth.gr} 
    \author[LAP]{Charikleia Meleti%
  %
  }
   \ead{meleti@auth.gr} 
      \affiliation[LAP]{
    organization={Laboratory of Atmospheric Physics, Physics Department,
Aristotle University of
Thessaloniki},city={Thessaloniki},postcode={54124},country={Greece},}
    \cortext[cor1]{Corresponding author}
    \fntext[1]{This is the first author footnote.}
    \fntext[2]{Another author footnote.}

  \begin{abstract}
  Here will be the abstract.
  \end{abstract}
    \begin{keyword}
    cloud enhancement \sep total solar radiation \sep global horizontal
irradiance \sep 
    over irradiance
  \end{keyword}

@@

~ ~

CE: cloud enhancement cases one minute

@@

{\centering \includegraphics[width=1\linewidth]{../images/P_CS_change-1}

@@

Our dataset, after the data selection processing, consists of 6144534
records of GHI measurements, of which \(1.764\,\%\) were identified as
CE events and \(0.036\,\%\) as ECE events. The highest recorded GHI due
to CE was \(1416.6\,\text{W}/\text{m}^2\) on 24~May 2007, and had the
maximum value of OIR at \(345.9\,\text{W}/\text{m}^2\). The relative
stronger CE event was \(49.7\,\%\) above the clear sky threshold, on
28~October 2016.

\hypertarget{long-term-trends}{%
\subsection{Long term trends}\label{long-term-trends}}

@@

{\centering \includegraphics[width=1\linewidth]{../images/P_energy-6}

@@

\caption{Trend of yearly cloud enhancement occurrences.}\label{fig:P-energy-N}
\end{figure}

@@

{\centering \includegraphics[width=1\linewidth]{../images/P_energy-5}

@@

\hypertarget{cloud-enhancements-climatology}{%
\subsection{Cloud enhancements
climatology}\label{cloud-enhancements-climatology}}

Another interesting aspect of the CE cases occurrences, is their
seasonal cycle. In
Figure\nobreakspace{}\ref{fig:relative-month-occurrences}, we have the
monthly box plot (whisker plot), where the values have been normalized
by the highest median value, that occurs on June. Although the number of
occurrences has a wide spread throughout the study period, the most
active period of CE occurrences is during May and June. During the
Winter (December -- February) the CE cases are about \(25\,\%\) of the
maximum. For the rest months, the occurrences gradually ramp between the
maximum and minimum. This seasonality is a combinations of the clouds
occurrences, and the relevant position of the sun during each season of
the year. Unfortunately, the lack of detailed data on cloud formation,
type and location is not allowing further analysis.

@@

{\centering \includegraphics[width=1\linewidth]{../images/clim_CE_month_norm_MAX_median_N-2}

@@

\caption{Seasonal statistics of the number of CE occurrences for each month, normalized to the maximum occurrences on June. The box represents the values of the low $25\,\%$ percentile to $75\,\%$ percentile, where the thick horizontal line inside is the mean, the vertical lines extend to the maximum and minimum values, the dots are outlier values, and the rhombus is the mean.}\label{fig:relative-month-occurrences}
\end{figure}

The distribution of the OIR intensity, follows an exponential decline
Figure\nobreakspace{}\ref{fig:ovir-distribution}, where there is an
inverse relation between the CE events frequency and their intensity.
This is expected, as the stronger the CE events become, the rarer the
particular atmospheric and sun conditions can occur. This fact, is
indicative to the magnitude and the probability of the expected OIRs
events over Thessaloniki. Similar distribution of CE occurrences have
been reported by \citet{Vamvakas2020}, where the magnitude of OIR were
higher due to the location of the city of Patras, \(2.5^\circ\) closer
to the equator.

@@

{\centering \includegraphics[width=1\linewidth]{../images/P-relative-distribution-diff-1}

@@

{\centering \includegraphics[width=1\linewidth]{../images/groups-1}

@@

{\centering \includegraphics[width=1\linewidth]{../images/P-groups-bin2d-1}

@@

\caption{Relation of the mean over irradiance and duration of GCE, where the color scale denotes the frequency of the respected events.}\label{fig:group-2d}
\end{figure}

\hypertarget{extreme-cloud-enhancement-events}{%
\subsection{Extreme cloud enhancement
events}\label{extreme-cloud-enhancement-events}}

An aspect of the CE events that is commonly reported and has some
significance on the solar energy production infrastructure are the
extreme CE events (ECE). Where solar irradiance exceeds the expected
irradiance on top of the atmosphere
(Equation\nobreakspace{}\ref{eq:ECE}). Analogous to
Figure\nobreakspace{}\ref{fig:relative-month-occurancies} we have
computed the distribution of the number of occurrences of ECE events by
month in Figure\nobreakspace{}\ref{fig:relative-month-occurancies-ECE}.
The most active period for ECE events is in the spring and the start of
the summer (March -- June), followed by a period in the late fall
(September and October). This is probably related to the weather
characteristics of these periods, where there are continuous
alternations between clear sky periods and clouds.

@@

{\centering \includegraphics[width=1\linewidth]{../images/clim_ECE_month_norm_MAX_median_N-2}

@@

\caption{Seasonal statistics of the number of ECE occurancies for each month, normalized to the maximum occurancies on March. The box represents the values of the low $25\,\%$ percentile to $75\,\%$ percentile, where the thick horizontal line inside is the mean, the vertical lines extend to the maximum and minimum vales, the dots are outlier values, and the rhombus is the mean.}\label{fig:relative-month-occurancies-ECE}
\end{figure}

The distribution of the ECE events
(Figure\nobreakspace{}\ref{fig:P-extreme-distribution}), shows that
there are rare cases where the OIR can exceed the TSI even more than
\(400\,\text{W}/\text{m}^2\), with the \(75\,\%\) of the cases to be
below \(200\,\text{W}/\text{m}^2\). Those finds are in accordance with
results form \citet{Vamvakas2020}, with the difference that the OIR
values are lower for Thessaloniki.

@@

{\centering \includegraphics[width=1\linewidth]{../images/P-extreme-distribution-1}

@@

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{../images/example_days-18}

@@

\caption{Example of cloud identification data for 2019-07-11. Where are denoted with red circles all the detected CE events, the cloud interaction with the GHI (blue crosses), and the thresholds of CE and the threshold of ECE events with red and black line respectively.}\label{fig:example-day}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{../images/example_years-13}

@@

\caption{Example of the relation of the measured GHI versus the reference clear sky irradiance for the year 2005, where the intensity of the detected over irradiance (OIR) is color coded.}\label{fig:example-year}
\end{figure}
